---
title: |
  | Mahcine Learning, Spring 2021
  | Homework 1
author: "Daniil Bulat, Jonas Josef Huwyler, Haochen Li,Giovanni Magagnin"
date: "March 3, 2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
---


```{r include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exerceise 2
**
Have a look at the file *lin_reg_iris_first_steps.R1* on StudyNet (Course Resources) and make sure you
understand all commands used in there. For comparability of the results please add the line *set.seed(123)*
in the beginning of the file. This will make the sample command return the same results every time you use
it.
**

The following is the basic setup for solving the exercise.
```{r}
# Set up
rm(list = ls())
set.seed(123)

# Shuffle the Iris data
Data = iris[sample(1:150),]

# and split into training and test data (80-20)
Data.Train = Data[1:120,]
Data.Test = Data[121:150,]
```

## Qestion 1: 

**
Determine the linear regression function in the form f(x1; x2) = m1x1 + m2x2 + c for predicting Sepal.Length depending on x1 =Petal.Length and x2 =Petal.Width on the training data.
**

The result below shows that the linear regression function for predicting would be 
$$Sepal.Length = 4.17 + 0.54*Petal.Length - 0.31*Petal.Width$$

```{r}
# Regression on the training data - Model 1: f(x1; x2) = m1x1 + m2x2 + c
Iris.Model1 = lm(Sepal.Length ~ Petal.Length + Petal.Width, data = Data.Train)
summary(Iris.Model1)
```

## Question 2:

**
Do the same for only one attribute, x1 =Petal.Length on the training data.
**

The result below shows that the linear regression function for predicting would be
$$Sepal.Length = 4.28 + 0.41*Petal.Length$$

```{r}
# Regression on the training data - Model 2: f(x1) = m1x1 + c
Iris.Model2 = lm(Sepal.Length ~ Petal.Length, data = Data.Train)
summary(Iris.Model2)

```

## Question 3:
**
Do the same for the three attributes, x1 =Petal.Length, x2 =Petal.Width, x3 =Sepal.Width on the training data.
**

The result below shows that the linear regression function for predicting would be 
$$Sepal.Length = \\
1.90 + 0.70*Petal.Length - 0.53*Petal.Width + 0.64*Sepal.Width$$
```{r}
# Regression on the training data - Model 3: f(x1; x2; x3) = m1x1 + m2x2 + m3x3 + c
Iris.Model3 = lm(Sepal.Length ~ Petal.Length + Petal.Width + Sepal.Width, data = Data.Train)
summary(Iris.Model3)
```

## Question 4:
**
Find the commands for mean and variance in R and compute the mean and the variance of Petal.Length and Petal.Width, respectively.
**

By using the commands from R, the mean of variance can be shown as following:

|              | Mean   | Variance |
|  :----:      | :----: |  :----:  |
| Petal.Length | 3.81   |   3.22   |
| Petal.Width  | 1.22   |   0.60   |


```{r}
mean(Data.Train$Petal.Length)
var(Data.Train$Petal.Length)
```

```{r}
mean(Data.Train$Petal.Width)
var(Data.Train$Petal.Width)
```

## Bonus Question:
**
Use the mean and variance commands to compute (or verify) the regression function for part 2) step by step without the lm-command, using the formula for simple linear regression.
**

Using the formula to solve the minimization problem of simple linear regression:
$$
\hat{\alpha} = \bar{y} -(\hat{\beta} \bar{x}) 
$$
$$
\hat{\beta} = \frac{\sum^n_{i=1}(x_i-\bar{x})(y_i-\bar{y})}{\sum^n_{i=1}(x_i-\bar{x})^2} = \frac{Cov(x,y)}{Var(x)}
$$
The regression coefficients calculated by mean and variance are identical to the lm-command:
$$Sepal.Length = 4.28 + 0.41*Petal.Length$$


```{r}
beta.hat = var(Data.Train$Sepal.Length,Data.Train$Petal.Length)/var(Data.Train$Petal.Length)
alpha.hat = mean(Data.Train$Sepal.Length) - beta.hat*mean(Data.Train$Petal.Length)
```

```{r}
alpha.hat
beta.hat
Iris.Model2$coefficients
```




