---
title: "Homework 1"
author: "Haochen Li"
date: "Mar 3, 2021"
output: 
  md_document:
    variant: markdown_github
---
# Exerceise 2
```{r include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Set up
rm(list = ls())
set.seed(123)

# Shuffle the Iris data
Data = iris[sample(1:150),]

# and split into training and test data (80-20)
Data.Train = Data[1:120,]
Data.Test = Data[121:150,]
```

## Qestion 1: 

Determine the Determine the linear regression function in the form f(x1; x2) = m1x1 + m2x2 + c for predicting Sepal.Length depending on x1 =Petal.Length and x2 =Petal.Width on the training data.

```{r}
# Regression on the training data - Model 1: f(x1; x2) = m1x1 + m2x2 + c
Iris.Model1 = lm(Sepal.Length ~ Petal.Length + Petal.Width, data = Data.Train)
summary(Iris.Model1)
```

## Question 2:

Do the same for only one attribute, x1 =Petal.Length on the training data.

```{r}
# Regression on the training data - Model 2: f(x1) = m1x1 + c
Iris.Model2 = lm(Sepal.Length ~ Petal.Length, data = Data.Train)
summary(Iris.Model2)

```

## Question 3:
Do the same for the three attributes, x1 =Petal.Length, x2 =Petal.Width, x3 =Sepal.Width on the training data.

```{r}
# Regression on the training data - Model 3: f(x1; x2; x3) = m1x1 + m2x2 + m3x3 + c
Iris.Model3 = lm(Sepal.Length ~ Petal.Length + Petal.Width + Sepal.Width, data = Data.Train)
summary(Iris.Model3)
```

## Question 4:

Find the commands for mean and variance in R and compute the mean and the variance of Petal.Length and Petal.Width, respectively.

```{r}
mean(Data.Train$Petal.Length)
var(Data.Train$Petal.Length)
```

```{r}
mean(Data.Train$Petal.Length)
var(Data.Train$Petal.Length)
```

## Bonus Question:
Use the mean and variance commands to compute (or verify) the regression function for part 2) step by step without the lm-command, using the formula for simple linear regression.

```{r}
beta.hat = var(Data.Train$Sepal.Length,Data.Train$Petal.Length)/var(Data.Train$Petal.Length)
alpha.hat = mean(Data.Train$Sepal.Length) - beta.hat*mean(Data.Train$Petal.Length)
```

```{r}
alpha.hat
beta.hat
Iris.Model2$coefficients
```




